Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-05 EconGrowth.Rnw
Now creating labels on figures and tables from knitr so that they can be referenced in the paper.

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(tikzDevice) #See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  echo=FALSE,    #Turns off code echo for all chunks. Set TRUE on an individual chunk to see it.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="The facts. Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies.">>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ),
                  ylim=list(c(1,10), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4)), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

############################
# Aggregates the Cobb-Douglas results into a big data table. No energy.
##
cobbDouglasParamsTableNoEnergy <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas (without energy) for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

############################
# Aggregates the Cobb-Douglas results into a big data table for the given energyType.
##
cobbDouglasParamsTableWithEnergy <- function(energyType){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow, energyType=energyType))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ",
                        " ", "$\\gamma$",  " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption=paste("Cobb-Douglas (with $", tolower(energyType), "$) for 1980-2011 (US, UK, JP) or 1991-2011 (others). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

@@

##############################
# Creates a graph that plots predicted GDP as lines and GDP data as open circles.
##
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the CD model without energy
  predictionsQ <- cobbDouglasPredictionsColumn("Q") #Predictions with Q
  predictionsX <- cobbDouglasPredictionsColumn("X") #Predictions with X
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictions, predictionsQ, predictionsX) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "red", "blue") #line and symbol colors
  symbols <- c(1, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP+predGDPQ+predGDPX ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "No energy", "With q", "With x")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ), 
                  ylim=list(c(1,8), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3)),
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
@

<<CD_Params_Table, results='asis', eval=FALSE>>=
print(cobbDouglasParamsTableNoEnergy(), caption.placement="top", sanitize.colnames.function = identity, size="\\tiny")
@

@@

\subsection{Cobb-Douglas with $Q$}

@@

@@

@@

@@

@@

>>>>

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(tikzDevice) #See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  warning=FALSE, #Tells whether to show warnings.
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=FALSE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

@@

Figure \ref{fig:Factors_Lattice_Graph} shows the ``facts'' as they are known.

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="The facts. Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ),
                  ylim=list(c(1,10), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4)), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

############################
# Aggregates the Cobb-Douglas results into a big data table. No energy.
##
cobbDouglasParamsTableNoEnergy <- function(){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption="Cobb-Douglas (without energy) for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", 
                    label="tab:CD_Parameters_No_Energy",
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

############################
# Aggregates the Cobb-Douglas results into a big data table for the given energyType.
##
cobbDouglasParamsTableWithEnergy <- function(energyType){
  #Do rbind on the results of creating a row in the table for every country abbreviation that we know.
  dataCD <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRow, energyType=energyType))
  colnames(dataCD) <- c(" ", "$\\lambda$", " ", 
                        " ", "$\\alpha$",  " ",
                        " ", "$\\beta$",   " ",
                        " ", "$\\gamma$",  " ")
  rownames(dataCD) <- countryAbbrevs
  tableCD <- xtable(dataCD, 
                    caption=paste("Cobb-Douglas (with $", tolower(energyType), "$) for 1980-2011 (US, UK, JP), 1991-2010 (CN and ZA), and 1991-2011 (SA, IR, TZ, and ZM). (Parameter estimates beneath symbol. 95\\% confidence interval bounds to left and right.)", sep=""), 
                    label=paste("tab:CD_Parameters_With_", energyType, sep=""),
                    digit = c(1, 4,4,4, 2,2,2, 2,2,2, 2,2,2),
                    align = "r|ccc|ccc|ccc|ccc") #Sets alignment of the numbers in the columns
  return(tableCD)
}

@@

##############################
# Creates a graph that plots predicted GDP as lines and GDP data as open circles.
##
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the CD model without energy
  predictionsQ <- cobbDouglasPredictionsColumn("Q") #Predictions with Q
  predictionsX <- cobbDouglasPredictionsColumn("X") #Predictions with X
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictions, predictionsQ, predictionsX) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "red", "blue") #line and symbol colors
  symbols <- c(1, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP+predGDPQ+predGDPX ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "No energy", "With $q$", "With $x$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ), 
                  ylim=list(c(1,8), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3)),
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
@

The Cobb-Douglas model without energy is given by

\begin{equation} \label{eq:CD_No_Energy}
  y = a^{\lambda (t-t_0)}k^{\alpha}l^{\beta}.
\end{equation}

Table \ref{tab:CD_Parameters_No_Energy} gives the parameters for the Cobb-Douglas model without energy.

<<CD_Params_Table, results='asis'>>=
print(cobbDouglasParamsTableNoEnergy(), caption.placement="top", sanitize.colnames.function = identity, size="\\tiny")
@

Figure \ref{fig:CD_Params_Graph} shows values and 95\% confidence intervals for the parameters for the Cobb-Douglas model (without energy).

@@

The Cobb-Douglas model with energy is given by

\begin{equation} \label{eq:CD_With_Energy}
  y = a^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma},
\end{equation}

\noindent where $e$ can be any of thermal energy ($q$), exergy ($x$), or useful work ($u$).

\subsection{Cobb-Douglas with $q$}

The Cobb-Douglas (with thermal energy, $q$) parameters are given in Table \ref{tab:CD_Parameters_With_Q}.

@@

Figure \ref{fig:CDq_Params_Graph} shows values and 95\% confidence intervals for the parameters for the Cobb-Douglas model (with $q$).

@@

The Cobb-Douglas (with exergy) parameters are given in Table \ref{tab:CD_Parameters_With_X}.

@@

Figure \ref{fig:CDx_Params_Graph} shows values and 95\% confidence intervals for the parameters for the Cobb-Douglas model (with $x$).

@@

\subsection{Cobb-Douglas Comparisons}

Figure \ref{fig:CD_GDP_Lattice_Graph} compares predictions from the Cobb-Douglas models (without energy, with $Q$, and with $x$) to historical data.

@@

\subsection{CES with $X$}

\subsection{CES with $U$}

\section{LINEX}

\subsection{LINEX with $Q$}

\subsection{LINEX with $X$}

\subsection{LINEX with $U$}
